Toward theories of deliberated preferences

### 2020-08-18 simple.tex
Dropped propsure and propposs

\RequirePackage[l2tabu, orthodox]{nag}
\documentclass[version=last, pagesize, twoside=off, bibliography=totoc, DIV=calc, fontsize=12pt, a4paper, french, english]{scrartcl}
\input{preamble/packages}
\input{preamble/math_basics}
\input{preamble/math_mine}
\input{preamble/redac}
\usepackage[normalem]{ulem}
%\input{preamble/draw}
%\input{preamble/acronyms}
\addtokomafont{labelinglabel}{\sffamily\bfseries}
\renewcommand{\phi}{\varphi}%TODO
\DeclareMathAlphabet{\mathup}{OT1}{\familydefault}{m}{n}

@@

\section{Normative theories}
\subsection{Definition}
\TODO{Consider using only $\phi$ being sure, and dropping everything related to possibility.}

One of the primitive elements of a normative theory is a set of arguments $\allargs$.
Given a set of arguments $\allargs$, define the set of finite sequences of elements in $\allargs$ (including the empty sequence) as $\allhist = \bigcup_{k \in \N} \allargs^{\intvl{1, k}}$.
Throughout the article, $\N$ includes zero and $\N^* = \N \setminus \set{0}$. Given $j, k \in \N$, the notation $\intvl{j, k}$ represents $\set{l \suchthat j ≤ l ≤ k}$. Note that $\intvl{1, 0} = \emptyset$, and given any set $S$, the notation $S^\emptyset$ represents the empty set. A generic element of $\allhist$ is denoted by $\hist$ and called a path.

Two supplementary primitive elements of a normative theory are its agenda $\allprops$, whose elements are called propositions; and a set of individuals $I$ that the theory applies to.
Given a set of individuals, a set of arguments and an agenda, an observation protocol ${\allleadsto}: I → \Phi^{\allhist}$ maps individuals and paths to propositions. 
Equivalently, given any individual $i \in I$, ${\ileadsto}: \allhist → \Phi$ maps paths to propositions for this individual. 
Given an individual $i \in I$, a path $\hist$ and a proposition $\phi$, the path is said to lead to the proposition for this individual iff $\hist \ileadsto \phi$.

A normative theory also defines an involution operator $¬$ on $\allprops$, thus a function $¬: \allprops → \allprops$ satisfying $\forall \prop \in \allprops: ¬¬\prop = \prop$.

A normative theory is a tuple $(\allprops, ¬, \allargs, I, {\allleadsto})$ 
%with $\allprops$ an agenda, $¬$ a negation operator on $\allprops$, $\allargs$ a set of arguments, $I$ a set of individuals, and ${\allleadsto}$ an observation protocol, 
satisfying the constraints described here above.
%where $\Phi = \allprops × \set{\mathit{poss}, \mathit{sure}}$.

This article also uses this symbol as a negation operator on $\Phi$, defined as follows, $\forall \prop \in \allprops$: $¬\propposs = (¬\prop, \mathit{sure})$ and $¬\propsure = (¬\prop, \mathit{poss})$. Observe that $¬¬\phi = \phi$. (The symbol $¬$ is also used in this article to negate classical mathematical propositions; this should create no confusion.)
As usual with binary relations, $\hist \nileadsto \phi$ means $(\hist, \phi) \notin {\ileadsto}$.

Given $i \in I, \phi \in \Phi, \hist \in \allhist$, say that $\hist$ leads strongly to $\phi$ iff $\hist \ileadsto \phi \land \hist \nileadsto ¬\phi$.
Define $\hist \ileadstost \phi$ iff $[\hist \text{ leads strongly to } \phi] \land [\phi = \propsure ⇒ \hist \text{ leads strongly to } \propposs]$.
(The visual proximity of the symbols $\ileadsto$ and $\ileadstost$ creates a small risk of confusion, but its usage seems justified by the fact that $\ileadsto$ is not used in the rest of this article, except in the proof of \cref{th:protcoh}.)

@@

Given $i \in I, \phi \in \Phi, \ar \in \allargs$, define $\ar \ileadstost \phi$ iff $\forall \hist \in \allhist \suchthat \ar \in \histend: \hist \ileadstost \phi$.

\begin{remark}
	I’d rather drop the negation and observe that if the protocol is defined so that $\hist \ileadsto \phi_1 ⇒ \hist \nileadsto \phi_2$, then $\phi_1 \in \iPhi ⇒ \phi_2 \notin \iPhi$. Observe that this is always possible. The definitions about completeness should use a negation operator in supplement to the normative theory tuple.
\end{remark}

\subsection{Classify propositions}
\begin{axiom}[Normative adequacy]
	\label{ax:norm}
	$\forall i \in I, \phi \in \Phi: 
		[(\exists \ar \in \allargs \suchthat \ar \ileadstost \phi) \land ¬(\exists \ar \in \allargs \suchthat \ar \ileadstost ¬\phi)] ⇒ \phi \in \iPhi \land ¬\phi \notin \iPhi.$
\end{axiom}

\Cref{ax:norm} is actually equivalent to the following property.
\begin{property}[Simple normative adequacy]
	\label{ax:snorm}
	$\forall i \in I, \phi \in \Phi: 
		[\exists \ar \in \allargs \suchthat \ar \ileadstost \phi] ⇒ \phi \in \iPhi \land ¬\phi \notin \iPhi.$
\end{property}

To see why this hold, consider the following definition and proposition.
\begin{definition}[Protocol coherence]
	\label{def:protcoh}
	$\forall i \in I, \phi \in \Phi$:
	\begin{equation}
		[\exists \ar \in \allargs \suchthat \ar \ileadstost \phi] ⇒ [\nexists \ar \in \allargs \suchthat \ar \ileadstost ¬\phi].
	\end{equation}
\end{definition}
\begin{proposition}
	\label{th:protcoh}
	Any normative theory satisfies Protocol coherence.
\end{proposition}
\begin{proof}
	Protocol coherence is satisfied by definition of $\ileadstost$. Indeed, if for some $\ar \in \allargs$, $\ar \ileadstost \phi$, then $\forall \ar_1 \in \allargs: (\ar_1, \ar) \nileadsto ¬\phi$; thus, for any $\ar' \in \allargs$, $\ar' \ileadstost ¬\phi$ is false, because it requires that $\forall \ar_1 \in \allargs: (\ar', \ar_1) \ileadsto ¬\phi$, and thus in particular, that $(\ar', \ar) \ileadsto ¬\phi$.
\end{proof}

@@

To further analyze this and related matters, given a normative theory and $i \in I$, say that a proposition $\phi \in \Phi$ is \emph{$i$-decidable} iff $\exists \ar \in \allargs \suchthat \ar \ileadstost \phi \lor \ar \ileadstost ¬\phi$.
The $i$-decidable propositions are those on which the normative theory permits to take a position. It will generally be unknown in practical applications, and may be empty.
The next property ensures an unambiguous interpretation of the content of $\iPhi$ for all $i$-decidable propositions. In this definition, $\oplus$ denotes the exclusive disjunction; thus $\phi \in \iPhi \oplus ¬\phi \in \iPhi$ is equivalent to: $\phi \in \iPhi ⇔ ¬\phi \notin \iPhi$. It can be seen to hold by using \cref{th:normeqsnorm}.

\begin{property}[Restricted interpretability]
	\label{def:restrinterpr}
	$\forall i \in I, \phi \in \Phi: [\exists \ar \in \allargs \suchthat \ar \ileadstost \phi \lor \ar \ileadstost ¬\phi] ⇒ [\phi \in \iPhi \oplus ¬\phi \in \iPhi]$.
\end{property}

\subsection{Non-completeness}
\label{sec:incompl}
\begin{definition}[Theory completeness]
	$\forall \phi \in \Phi: [\exists \ar \in \allargs \suchthat \ar \ileadstost \phi] \lor [\exists \ar \in \allargs \suchthat \ar \ileadstost ¬\phi]$.
\end{definition}
\begin{definition}[Unrestricted interpretability]
	$\forall \phi \in \Phi: \phi \in \iPhi \oplus ¬\phi \in \iPhi$.
\end{definition}
\begin{definition}[Deliberation completeness]
	$\forall \phi \in \Phi: \phi \in \iPhi \lor ¬\phi \in \iPhi$.
\end{definition}
The following proposition directly follows from the definitions, together with \cref{th:protcoh}.
\begin{proposition}
	1) Under Normative adequacy, Theory completeness implies Deliberation completeness. 2) Unrestricted interpretability implies Deliberation completeness.
\end{proposition}
Theories as defined here do not mandate Theory completeness. This is good for two reasons. First, even the weaker Deliberation completeness can be considered too strong. Second, it is dubious that, for non trivial decision problems, theories can be found for which both Normative adequacy is normatively compelling and Theory completeness holds.

@@

An initial argumentation relation ${\gleadsto} \subseteq \allargs × \allprops$ relates some arguments to some propositions. Its role in an empirical theory is to define the propositions that the empirical theory supports (corresponding to $\gPhie$, the range of ${\gleadsto}$) and, for each such proposition, to define the arguments that the theory wants to use as initial arguments to argue in favor of this proposition. The set of initial arguments is $\gleadstoinv(\Phi) = \set{\ar \in \allargs \suchthat \exists \phi \in \Phi \suchthat \ar \gleadsto \phi}$.

@@

This mechanism permits to define defenders of defenders of an argument, $\dfp(\dfp(\ar))$, defenders of defenders of defenders of an argument, … An empirical theory must however be a priori restricted in the maximal “depth” of its argumentation strategy, in the following sense.
Given any set $S$ and function $f: S → S$, let $f^0$ represent the identity function, and given $k \in \N$, let $f^{k + 1} = f \circ f^k$.
An empirical theory must satisfy
$\exists k \in \N \suchthat (\dfp)^k(\gleadstoinv(\Phi)) = \emptyset$.
This does not imply that the relations $\gbeats$ or $\dbeats$ should be finite, however.

@@

An example permits to ensure that the constraint of finite depth does not mandate finiteness of the relations $\gbeats$ or $\dbeats$.
\begin{proposition}[Possibility of infinite relations]
	An empirical theory may involve infinite relations $\gbeats$ and $\dbeats$.
\end{proposition}
\begin{proof}
	Consider an agenda containing two propositions, $\allprops = \set{\phi, ¬\phi}$, an infinite set of arguments $\allargs = \set{\ar_k \suchthat k \in \N}$ with $\forall j, k \in \N: \ar_j ≠ \ar_k$, and consider any set of individuals $I$ and observation protocol $\allleadsto$. 
	Define $2\N^* = \set{k \suchthat k / 2 \in \N^*}$ as the set of non-zero even integers.
	Define ${\gleadsto} = \set{(\ar_0, \phi)}$.
	Define ${\gbeats} = \set{(\ar_{k - 1}, \ar_0) \suchthat k \in 2\N^*}$ so that all odd numbered arguments are planned attacks against $\ar_0$. 
	Consider an infinite set of fragments $\alldargs = \set{\ard[k] \suchthat k \in 2\N^*}$ with $\forall j, k \in 2\N^*: \ard[j] ≠ \ard[k]$ so that fragments correspond to non-zero even integers, and define ${\dbeats} = \set{(\ard[k], \ar_{k - 1}) \suchthat k \in 2\N^*}$ so that each (even numbered) defense fragment attacks exactly one (odd numbered) planned attacker of $\ar_0$. 
	Given any coalition of fragments $\dargs$ defending $\ar_0$, consider the set of indices corresponding to the fragments in the coalition, $K_{\dargs} = \set{k \in 2\N^* \suchthat \exists \ard[k] \in \dargs}$; 
	define $j_{\dargs} = \sum_{k \in K_{\dargs}}$ as the sum of these indices;
	and define $\ar_0 + \dargs = \ar_{j_{\dargs}}$. 
%	$\ard[k] \in \alldargs$, define $\ar_0 + \set{\ard[k]} = \ar_k$. 
	Consider an empirical theory $(\gleadsto, \gbeats, \alldargs, \dbeats, +)$ for the normative theory $(\allprops, ¬, \allargs, I, {\allleadsto})$. That this is indeed a legal empirical theory can be confirmed by checking that all constraints are satisfied, in particular, $(\dfp)^2(\gleadstoinv(\Phi)) = \dfp(\dfp(\ar_0)) = \dfp(\set{\ar_k \suchthat k \in 2\N^*}) = \emptyset$ because there is no planned attack against any argument in $\set{\ar_k \suchthat k \in 2\N^*}$. This theory involves infinite relations $\gbeats$ and $\dbeats$.
\end{proof}

@@

Given $i \in I, \phi \in \Phi, \ar \in \allargs$ and $\args \subseteq \gbeatsinv(\ar)$, define $\args \ibeats \ar$ iff $\exists \hist \in \allhist \suchthat \args = \gbeatsinv(\ar) \cap \hist \land \ar \in \histend \land \hist \nileadstost \phi$. 
Such a set $\args$, when not empty, represents an effective set of planned attackers of $\ar$ against $\phi$. 
The notation uses a striked through $\phi$ in superscript to suggest that when $\args \ibeats \ar$, the coalition $\args$ argues against $\phi$, from the point of view of $i$.
I write that $\ar$ \emph{$i$-defends} $\phi$ to mean that $\emptyset \nibeats \ar$.

Given $i \in I, \phi \in \gPhie, \ar \in \allargs$, define $\di(\ar) \subseteq \powersetz{\alldargs}$ as $\di(\ar) = \set{\dbeatsinv(\args) \suchthat \emptyset ≠ \args \subseteq \gbeatsinv(\ar) \land \args \ibeats \ar}$. It represents the coalitions of fragments defending $\ar$ (when arguing for $\phi$), from the point of view of $i$, against its effective planned attackers. 
The following proposition shows that these coalitions can be added to $\ar$, and will be useful for other results.
\begin{proposition}
	\label{th:indf}
	$\forall i \in I, \phi \in \gPhie, \ar \in \allargs: \dargs \in \di(\ar) ⇒ \emptyset ≠ \dargs \subseteq \df(\ar)$.
\end{proposition}
\begin{proof}
	I will prove that $\forall \ar \in \allargs: \set{\dbeatsinv(\args) \suchthat \emptyset ≠ \args \subseteq \gbeatsinv(\ar)} = \powersetz{\df(\ar)}$, from which the conclusion follows, and which can be useful to clarify the relationship between $\di(\ar)$ and $\df(\ar)$. Note that by definition of $\df(\ar)$, $\dargs \in \powersetz{\df(\ar)}$ iff $\emptyset ≠ \dargs \subseteq \dbeatsinv(\gbeatsinv(\ar))$. 
	Thus, we have to prove that
	$\forall \dargs \subseteq \allargs$:
	\begin{equation}
		\exists \emptyset ≠ \args \subseteq \gbeatsinv(\ar) \suchthat \dbeatsinv(\args) = \dargs ⇔ \emptyset ≠ \dargs \subseteq \dbeatsinv(\gbeatsinv(\ar)).
	\end{equation}

@@

Given $\ar \in \allargs$, define $\dip(\ar) = \set{(\ar + \dargs) \suchthat \dargs \in \di(\ar)}$, representing the subset of defenders of $\ar$ required for convincing $i$ of $\phi$.
Given $i \in I, \phi \in \gPhie$, define $\gargs \subseteq \allargs$ as the transitive closure of ${\gleadstoinv}(\phi)$ under $\dip$: $\gargs = \cup_{k \in \N} (\dip)^k({\gleadstoinv}(\phi))$.

\begin{definition}[Convincingness]
	An empirical theory is \emph{convincing} iff $\forall i \in I, \phi \in \gPhie, \ar \in \gargs: \ar$ $i$-defends $\phi$.
\end{definition}

A \emph{falsification instance} of an empirical theory is a tuple $(i, \phi, \hist, \ar) \in I × \gPhie × \allhist × \gargs$ such that $\gbeatsinv(\ar) \cap \hist = \emptyset \land \ar \in \histend \land \hist \nileadstost \phi$. 
I also write that $(i, \phi, \hist, \ar)$ falsifies $\gamma$ to mean that it is a falsification instance of $\gamma$.

\begin{theorem}
	An empirical theory is convincing iff it has no falsification instances.
\end{theorem}
\begin{proof}
	For all $i \in I, \phi \in \gPhie, \ar \in \gargs$: $\ar$ does not $i$-defend $\phi$ iff $\exists \hist \in \allhist \suchthat \left[\gbeatsinv(\ar) \cap \hist = \emptyset \land \ar \in \histend \land \hist \nileadstost \phi\right]$ iff $\exists \hist \in \allhist$ such that $(i, \phi, \hist, \ar)$ falsifies $\gamma$. Therefore, $\gamma$ is not convincing iff $\exists (i, \phi, \ar) \suchthat \ar$ does not $i$-defend $\phi$ iff $\exists \hist \suchthat (i, \phi, \hist, \ar)$ falsifies $\gamma$ iff $\gamma$ has a falsification instance.
\end{proof}

\begin{definition}[Validity]
	An empirical theory is \emph{valid} iff $\forall i \in I, \phi \in \gPhie: [\phi \in \iPhi \land ¬\phi \notin \iPhi]$.
\end{definition}

\begin{lemma}
	\label{th:produce}
	Given $\phi \in \gPhie$, $\ar_0 \in \gargs$ that $i$-defends $\phi$, and $\hist \nileadstost \phi$ with $\ar_0 \in \histend$, define $\args =  \gbeatsinv(\ar_0) \cap \hist$, $\dargs = \dbeatsinv(\args)$, and $\ar_2 = \ar_0 + \dargs$. Then, $\ar_2 \in \gargs \cap \dfp(\ar_0)$.
\end{lemma}
\begin{proof}
	By definition of $\ibeats$, $\args \ibeats \ar_0$.
	Also, because $\ar_0$ $i$-defends $\phi$, $\args ≠ \emptyset$. 
	Therefore, $\dargs \in \di(\ar_0)$.
	It follows that $\dargs \in \df(\ar_0)$, by \cref{th:indf}, and that $\ar_2 \in \dfp(\ar_0)$, by definition of $\dfp$. It also follows that $\ar_2 \in \dip(\ar_0)$, hence, that $\ar_2 \in \gargs$.
\end{proof}

\begin{theorem}
	A normatively adequate and convincing empirical theory is valid.
\end{theorem}
\begin{proof}
	Given any $i \in I$ and $\phi \in \gPhie$, suffices to show that for some $\ar \in \allargs$, $\ar \ileadstost \phi$; the conclusion then follows thanks to simple normative adequacy (that \cref{th:normeqsnorm} shows to be equivalent to normative adequacy).

>>>>

\RequirePackage[l2tabu, orthodox]{nag}
\documentclass[version=last, pagesize, twoside=off, bibliography=totoc, DIV=calc, fontsize=12pt, a4paper, french, english]{scrartcl}
\input{preamble/packages}
\input{preamble/math_basics}
\input{preamble/math_mine}
\input{preamble/redac}
\usepackage[normalem]{ulem}
%\input{preamble/draw}
%\input{preamble/acronyms}
\addtokomafont{labelinglabel}{\sffamily\bfseries}
\DeclareMathAlphabet{\mathup}{OT1}{\familydefault}{m}{n}

@@

\section{Normative theories}
\subsection{Definition}
One of the primitive elements of a normative theory is a set of arguments $\allargs$.
(The calligraphic letter $S$ is used to denote the arguments in order to reserve the letter $A$ for sets of alternatives.)
Given a set of arguments $\allargs$, define the set of finite sequences of elements in $\allargs$ (including the empty sequence) as $\allhist = \bigcup_{k \in \N} \allargs^{\intvl{1, k}}$.
Throughout the article, $\N$ includes zero and $\N^* = \N \setminus \set{0}$. Given $j, k \in \N$, the notation $\intvl{j, k}$ represents $\set{l \suchthat j ≤ l ≤ k}$. Note that $\intvl{1, 0} = \emptyset$, and given any set $S$, the notation $S^\emptyset$ represents the empty set. A generic element of $\allhist$ is denoted by $\hist$ and called a path.

Two supplementary primitive elements of a normative theory are its agenda $\allprops$, whose elements are called propositions; and a set of individuals $I$ that the theory applies to.
Given a set of individuals, a set of arguments and an agenda, an observation protocol ${\allleadsto}: I → \allprops^{\allhist}$ maps individuals and paths to propositions. 
Equivalently, given any individual $i \in I$, ${\ileadsto}: \allhist → \allprops$ maps paths to propositions for this individual. 
Given an individual $i \in I$, a path $\hist$ and a proposition $\prop$, the path is said to lead to the proposition for this individual iff $\hist \ileadsto \prop$.

A normative theory also defines an involution operator $¬$ on $\allprops$, thus a function $¬: \allprops → \allprops$ satisfying $\forall \prop \in \allprops: ¬¬\prop = \prop$.
(The symbol $¬$ is also used in this article to negate classical mathematical propositions; this should create no confusion.)

A normative theory is a tuple $(\allprops, ¬, \allargs, I, {\allleadsto})$ 
%with $\allprops$ an agenda, $¬$ a negation operator on $\allprops$, $\allargs$ a set of arguments, $I$ a set of individuals, and ${\allleadsto}$ an observation protocol, 
satisfying the constraints described here above.
%where $\allprops = \allprops × \set{\mathit{poss}, \mathit{sure}}$.

As usual with binary relations, $\hist \nileadsto \prop$ means $(\hist, \prop) \notin {\ileadsto}$.

Given $i \in I, \prop \in \allprops, \hist \in \allhist$, say that $\hist$ leads strongly to $\prop$, $\hist \ileadstost \prop$, iff $\hist \ileadsto \prop \land \hist \nileadsto ¬\prop$.
(The visual proximity of the symbols $\ileadsto$ and $\ileadstost$ creates a small risk of confusion, but its usage seems justified by the fact that $\ileadsto$ is not used in the rest of this article, except in the proof of \cref{th:protcoh}.)

@@

Given $i \in I, \prop \in \allprops, \ar \in \allargs$, define $\ar \ileadstost \prop$ iff $\forall \hist \in \allhist \suchthat \ar \in \histend: \hist \ileadstost \prop$.

\begin{remark}
	I’d rather drop the negation and observe that if the protocol is defined so that $\hist \ileadsto \prop_1 ⇒ \hist \nileadsto \prop_2$, then $\prop_1 \in \iprops ⇒ \prop_2 \notin \iprops$. Observe that this is always possible. The definitions about completeness should use a negation operator in supplement to the normative theory tuple.
\end{remark}

\subsection{Classify propositions}
\begin{axiom}[Normative adequacy]
	\label{ax:norm}
	$\forall i \in I, \prop \in \allprops: 
		[(\exists \ar \in \allargs \suchthat \ar \ileadstost \prop) \land ¬(\exists \ar \in \allargs \suchthat \ar \ileadstost ¬\prop)] ⇒ \prop \in \iprops \land ¬\prop \notin \iprops.$
\end{axiom}

\Cref{ax:norm} is actually equivalent to the following property.
\begin{property}[Simple normative adequacy]
	\label{ax:snorm}
	$\forall i \in I, \prop \in \allprops: 
		[\exists \ar \in \allargs \suchthat \ar \ileadstost \prop] ⇒ \prop \in \iprops \land ¬\prop \notin \iprops.$
\end{property}

To see why this hold, consider the following definition and proposition.
\begin{definition}[Protocol coherence]
	\label{def:protcoh}
	$\forall i \in I, \prop \in \allprops$:
	\begin{equation}
		[\exists \ar \in \allargs \suchthat \ar \ileadstost \prop] ⇒ [\nexists \ar \in \allargs \suchthat \ar \ileadstost ¬\prop].
	\end{equation}
\end{definition}
\begin{proposition}
	\label{th:protcoh}
	Any normative theory satisfies Protocol coherence.
\end{proposition}
\begin{proof}
	Protocol coherence is satisfied by definition of $\ileadstost$. Indeed, if for some $\ar \in \allargs$, $\ar \ileadstost \prop$, then $\forall \ar_1 \in \allargs: (\ar_1, \ar) \nileadsto ¬\prop$; thus, for any $\ar' \in \allargs$, $\ar' \ileadstost ¬\prop$ is false, because it requires that $\forall \ar_1 \in \allargs: (\ar', \ar_1) \ileadsto ¬\prop$, and thus in particular, that $(\ar', \ar) \ileadsto ¬\prop$.
\end{proof}

@@

To further analyze this and related matters, given a normative theory and $i \in I$, say that a proposition $\prop \in \allprops$ is \emph{$i$-decidable} iff $\exists \ar \in \allargs \suchthat \ar \ileadstost \prop \lor \ar \ileadstost ¬\prop$.
The $i$-decidable propositions are those on which the normative theory permits to take a position. It will generally be unknown in practical applications, and may be empty.
The next property ensures an unambiguous interpretation of the content of $\iprops$ for all $i$-decidable propositions. In this definition, $\oplus$ denotes the exclusive disjunction; thus $\prop \in \iprops \oplus ¬\prop \in \iprops$ is equivalent to: $\prop \in \iprops ⇔ ¬\prop \notin \iprops$. It can be seen to hold by using \cref{th:normeqsnorm}.

\begin{property}[Restricted interpretability]
	\label{def:restrinterpr}
	$\forall i \in I, \prop \in \allprops: [\exists \ar \in \allargs \suchthat \ar \ileadstost \prop \lor \ar \ileadstost ¬\prop] ⇒ [\prop \in \iprops \oplus ¬\prop \in \iprops]$.
\end{property}

\subsection{Non-completeness}
\label{sec:incompl}
\begin{definition}[Theory completeness]
	$\forall \prop \in \allprops: [\exists \ar \in \allargs \suchthat \ar \ileadstost \prop] \lor [\exists \ar \in \allargs \suchthat \ar \ileadstost ¬\prop]$.
\end{definition}
\begin{definition}[Unrestricted interpretability]
	$\forall \prop \in \allprops: \prop \in \iprops \oplus ¬\prop \in \iprops$.
\end{definition}
\begin{definition}[Deliberation completeness]
	$\forall \prop \in \allprops: \prop \in \iprops \lor ¬\prop \in \iprops$.
\end{definition}
The following proposition directly follows from the definitions, together with \cref{th:protcoh}.
\begin{proposition}
	1) Under Normative adequacy, Theory completeness implies Deliberation completeness. 2) Unrestricted interpretability implies Deliberation completeness.
\end{proposition}
Theories as defined here do not mandate Theory completeness. This is good for two reasons. First, even the weaker Deliberation completeness can be considered too strong. Second, it is dubious that, for non trivial decision problems, theories can be found for which both Normative adequacy is normatively compelling and Theory completeness holds.

@@

An initial argumentation relation ${\gleadsto} \subseteq \allargs × \allprops$ relates some arguments to some propositions. Its role in an empirical theory is to define the propositions that the empirical theory supports (corresponding to $\gpropse$, the range of ${\gleadsto}$) and, for each such proposition, to define the arguments that the theory wants to use as initial arguments to argue in favor of this proposition. The set of initial arguments is $\gleadstoinv(\allprops) = \set{\ar \in \allargs \suchthat \exists \prop \in \allprops \suchthat \ar \gleadsto \prop}$.

@@

This mechanism permits to define defenders of defenders of an argument, $\dfp(\dfp(\ar))$, defenders of defenders of defenders of an argument, … An empirical theory must however be a priori restricted in the maximal “depth” of its argumentation strategy, in the following sense.
Given any set $S$ and function $f: S → S$, let $f^0$ represent the identity function, and given $k \in \N$, let $f^{k + 1} = f \circ f^k$.
An empirical theory must satisfy
$\exists k \in \N \suchthat (\dfp)^k(\gleadstoinv(\allprops)) = \emptyset$.
This does not imply that the relations $\gbeats$ or $\dbeats$ should be finite, however.

@@

An example permits to ensure that the constraint of finite depth does not mandate finiteness of the relations $\gbeats$ or $\dbeats$.
\begin{proposition}[Possibility of infinite relations]
	An empirical theory may involve infinite relations $\gbeats$ and $\dbeats$.
\end{proposition}
\begin{proof}
	Consider an agenda containing two propositions, $\allprops = \set{\prop, ¬\prop}$, an infinite set of arguments $\allargs = \set{\ar_k \suchthat k \in \N}$ with $\forall j, k \in \N: \ar_j ≠ \ar_k$, and consider any set of individuals $I$ and observation protocol $\allleadsto$. 
	Define $2\N^* = \set{k \suchthat k / 2 \in \N^*}$ as the set of non-zero even integers.
	Define ${\gleadsto} = \set{(\ar_0, \prop)}$.
	Define ${\gbeats} = \set{(\ar_{k - 1}, \ar_0) \suchthat k \in 2\N^*}$ so that all odd numbered arguments are planned attacks against $\ar_0$. 
	Consider an infinite set of fragments $\alldargs = \set{\ard[k] \suchthat k \in 2\N^*}$ with $\forall j, k \in 2\N^*: \ard[j] ≠ \ard[k]$ so that fragments correspond to non-zero even integers, and define ${\dbeats} = \set{(\ard[k], \ar_{k - 1}) \suchthat k \in 2\N^*}$ so that each (even numbered) defense fragment attacks exactly one (odd numbered) planned attacker of $\ar_0$. 
	Given any coalition of fragments $\dargs$ defending $\ar_0$, consider the set of indices corresponding to the fragments in the coalition, $K_{\dargs} = \set{k \in 2\N^* \suchthat \exists \ard[k] \in \dargs}$; 
	define $j_{\dargs} = \sum_{k \in K_{\dargs}}$ as the sum of these indices;
	and define $\ar_0 + \dargs = \ar_{j_{\dargs}}$. 
%	$\ard[k] \in \alldargs$, define $\ar_0 + \set{\ard[k]} = \ar_k$. 
	Consider an empirical theory $(\gleadsto, \gbeats, \alldargs, \dbeats, +)$ for the normative theory $(\allprops, ¬, \allargs, I, {\allleadsto})$. That this is indeed a legal empirical theory can be confirmed by checking that all constraints are satisfied, in particular, $(\dfp)^2(\gleadstoinv(\allprops)) = \dfp(\dfp(\ar_0)) = \dfp(\set{\ar_k \suchthat k \in 2\N^*}) = \emptyset$ because there is no planned attack against any argument in $\set{\ar_k \suchthat k \in 2\N^*}$. This theory involves infinite relations $\gbeats$ and $\dbeats$.
\end{proof}

@@

Given $i \in I, \prop \in \allprops, \ar \in \allargs$ and $\args \subseteq \gbeatsinv(\ar)$, define $\args \ibeats \ar$ iff $\exists \hist \in \allhist \suchthat \args = \gbeatsinv(\ar) \cap \hist \land \ar \in \histend \land \hist \nileadstost \prop$. 
Such a set $\args$, when not empty, represents an effective set of planned attackers of $\ar$ against $\prop$. 
The notation uses a striked through $\prop$ in superscript to suggest that when $\args \ibeats \ar$, the coalition $\args$ argues against $\prop$, from the point of view of $i$.
I write that $\ar$ \emph{$i$-defends} $\prop$ to mean that $\emptyset \nibeats \ar$.

Given $i \in I, \prop \in \gpropse, \ar \in \allargs$, define $\di(\ar) \subseteq \powersetz{\alldargs}$ as $\di(\ar) = \set{\dbeatsinv(\args) \suchthat \emptyset ≠ \args \subseteq \gbeatsinv(\ar) \land \args \ibeats \ar}$. It represents the coalitions of fragments defending $\ar$ (when arguing for $\prop$), from the point of view of $i$, against its effective planned attackers. 
The following proposition shows that these coalitions can be added to $\ar$, and will be useful for other results.
\begin{proposition}
	\label{th:indf}
	$\forall i \in I, \prop \in \gpropse, \ar \in \allargs: \dargs \in \di(\ar) ⇒ \emptyset ≠ \dargs \subseteq \df(\ar)$.
\end{proposition}
\begin{proof}
	I will prove that $\forall \ar \in \allargs: \set{\dbeatsinv(\args) \suchthat \emptyset ≠ \args \subseteq \gbeatsinv(\ar)} = \powersetz{\df(\ar)}$, from which the conclusion follows, and which can be useful to clarify the relationship between $\di(\ar)$ and $\df(\ar)$. Note that by definition of $\df(\ar)$, $\dargs \in \powersetz{\df(\ar)}$ iff $\emptyset ≠ \dargs \subseteq \dbeatsinv(\gbeatsinv(\ar))$. 
	Thus, we have to prove that
	$\forall \dargs \subseteq \allargs$:
	\begin{equation}
		\exists \emptyset ≠ \args \subseteq \gbeatsinv(\ar) \suchthat \dbeatsinv(\args) = \dargs ⇔ \emptyset ≠ \dargs \subseteq \dbeatsinv(\gbeatsinv(\ar)).
	\end{equation}

@@

Given $\ar \in \allargs$, define $\dip(\ar) = \set{(\ar + \dargs) \suchthat \dargs \in \di(\ar)}$, representing the subset of defenders of $\ar$ required for convincing $i$ of $\prop$.
Given $i \in I, \prop \in \gpropse$, define $\gargs \subseteq \allargs$ as the transitive closure of ${\gleadstoinv}(\prop)$ under $\dip$: $\gargs = \cup_{k \in \N} (\dip)^k({\gleadstoinv}(\prop))$.

\begin{definition}[Convincingness]
	An empirical theory is \emph{convincing} iff $\forall i \in I, \prop \in \gpropse, \ar \in \gargs: \ar$ $i$-defends $\prop$.
\end{definition}

A \emph{falsification instance} of an empirical theory is a tuple $(i, \prop, \hist, \ar) \in I × \gpropse × \allhist × \gargs$ such that $\gbeatsinv(\ar) \cap \hist = \emptyset \land \ar \in \histend \land \hist \nileadstost \prop$. 
I also write that $(i, \prop, \hist, \ar)$ falsifies $\gamma$ to mean that it is a falsification instance of $\gamma$.

\begin{theorem}
	An empirical theory is convincing iff it has no falsification instances.
\end{theorem}
\begin{proof}
	For all $i \in I, \prop \in \gpropse, \ar \in \gargs$: $\ar$ does not $i$-defend $\prop$ iff $\exists \hist \in \allhist \suchthat \left[\gbeatsinv(\ar) \cap \hist = \emptyset \land \ar \in \histend \land \hist \nileadstost \prop\right]$ iff $\exists \hist \in \allhist$ such that $(i, \prop, \hist, \ar)$ falsifies $\gamma$. Therefore, $\gamma$ is not convincing iff $\exists (i, \prop, \ar) \suchthat \ar$ does not $i$-defend $\prop$ iff $\exists \hist \suchthat (i, \prop, \hist, \ar)$ falsifies $\gamma$ iff $\gamma$ has a falsification instance.
\end{proof}

\begin{definition}[Validity]
	An empirical theory is \emph{valid} iff $\forall i \in I, \prop \in \gpropse: [\prop \in \iprops \land ¬\prop \notin \iprops]$.
\end{definition}

\begin{lemma}
	\label{th:produce}
	Given $\prop \in \gpropse$, $\ar_0 \in \gargs$ that $i$-defends $\prop$, and $\hist \nileadstost \prop$ with $\ar_0 \in \histend$, define $\args =  \gbeatsinv(\ar_0) \cap \hist$, $\dargs = \dbeatsinv(\args)$, and $\ar_2 = \ar_0 + \dargs$. Then, $\ar_2 \in \gargs \cap \dfp(\ar_0)$.
\end{lemma}
\begin{proof}
	By definition of $\ibeats$, $\args \ibeats \ar_0$.
	Also, because $\ar_0$ $i$-defends $\prop$, $\args ≠ \emptyset$. 
	Therefore, $\dargs \in \di(\ar_0)$.
	It follows that $\dargs \in \df(\ar_0)$, by \cref{th:indf}, and that $\ar_2 \in \dfp(\ar_0)$, by definition of $\dfp$. It also follows that $\ar_2 \in \dip(\ar_0)$, hence, that $\ar_2 \in \gargs$.
\end{proof}

\begin{theorem}
	A normatively adequate and convincing empirical theory is valid.
\end{theorem}
\begin{proof}
	Given any $i \in I$ and $\prop \in \gpropse$, suffices to show that for some $\ar \in \allargs$, $\ar \ileadstost \prop$; the conclusion then follows thanks to simple normative adequacy (that \cref{th:normeqsnorm} shows to be equivalent to normative adequacy).
